---
title: 'cfDNA Project: CCF Plots Report (DNA Presence)'
author: 'Pierre Levy'
date: '`r Sys.Date()`'
output:
    html_document:
        toc: true
        toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE
)
library(tidyverse)
library(data.table)
library(DT)
library(viridis)
library(paletteer)
library(ggridges)
library(ggbeeswarm)
library(patchwork)
library(ggrepel)
library(plotly)
library(ggbreak)
library(ggVennDiagram)

setwd("/Users/plevy/Library/CloudStorage/OneDrive-VHIO/labagros/pierre/proj/collab/andrea/CCF_plots/dna_presence")
```

```{r}
# Specify the directory path
samples <- c(
    "GOIGROS001",
    "GOIGROS002",
    "GOIGROS003",
    "GOIGROS004",
    "Motricolor002",
    "GOIGROS008",
    "TIL012",
    "TIL015"
)

directory_path <- "/Users/plevy/Library/CloudStorage/OneDrive-VHIO/labagros/pierre/proj/collab/andrea"

# Find all CCF.annot.tsv files in the specified directory for each sample
files <- list()
for (sample in samples) {
    ccf_table_path <- list.files(
        paste0(directory_path, "/", sample, "/min_callers_indels_2"),
        full.names = TRUE,
        pattern = "CCF.annot.tsv$"
    )

    files[[sample]] <- ccf_table_path
}

# Find all tested epitope files
tested_epitopes_files <- list()
for (sample in samples) {
    tested_path <- list.files(
        paste0(directory_path, "/", sample),
        full.names = TRUE,
        pattern = "tested_epitopes.txt"
    )

    tested_epitopes_files[[sample]] <- tested_path
}

# Find all epitope variant key files
epitope_variant_key_files <- list()
for (sample in samples) {
    key_path <- list.files(
        paste0(directory_path, "/", sample),
        full.names = TRUE,
        pattern = "epitope_variant_key.tsv"
    )

    epitope_variant_key_files[[sample]] <- key_path
}
```

```{r, echo = TRUE}
# POST VARIANT CALLING FILTERS
# TVAF >=3%
# Variant reads >= 4
# Coverage >= 9 in respective tumor (FrTu OR cfDNA) and normal samples for all plots except for analysis involving CCF data (Coverage >= 9 in both tumor and normal samples (FrTu AND cfDNA AND Normal))
# Called by 2 or more callers both SNV and INDELS
# All mutations that normal VAF = 0
# If normal has VAF, VAF >5X in tumor compared to normal

# FILTERS FOR DNA PRESENCE
# Minimum VAF and minimum ALT counts
minVaf <- 0.01
minAltReads <- 1
# OR minimum VAF only
minVaf_only <- 0.01

# For CCF analysis
# minimum coverage for all 3 sample types
min_cov <- 9
```

```{r}
# Function to create table with data from all each sample
get_ccf <- function(x, y) { # The y is necessary for the purrr::imap function later to access the names of the elements of the named list
    ccf_tbl <- fread(x)
    # Change NA values in vaf_FrTu and vaf_cfDNA to 0
    # NA for these vafs if no reads at all for REF and ALT for position in FrTu or cfDNA sample (especially for TIL patients as it was not a criteria for calling to have min coverage in all samples)
    ccf_tbl$vaf_FrTu[is.na(ccf_tbl$vaf_FrTu)] <- 0
    ccf_tbl$vaf_cfDNA[is.na(ccf_tbl$vaf_cfDNA)] <- 0
    ccf_tbl <- ccf_tbl %>%
        mutate(
            sample = ifelse(y == "Motricolor002", "GOIGROS005", y),
            dna_presence =
                ifelse(((vaf_FrTu >= minVaf & vaf_cfDNA >= minVaf & ALT_counts_FrTu >= minAltReads & ALT_counts_cfDNA >= minAltReads) | (vaf_FrTu >= minVaf_only & vaf_cfDNA >= minVaf_only)), "Shared",
                    ifelse((((vaf_FrTu >= minVaf & ALT_counts_FrTu >= minAltReads) & (vaf_cfDNA < minVaf | ALT_counts_cfDNA < minAltReads)) | (vaf_FrTu >= minVaf_only & vaf_cfDNA < minVaf_only)), "FrTu_only",
                        ifelse((((vaf_FrTu < minVaf | ALT_counts_FrTu < minAltReads) & (vaf_cfDNA >= minVaf & ALT_counts_cfDNA >= minAltReads)) | (vaf_cfDNA >= minVaf_only & vaf_FrTu < minVaf_only)), "cfDNA_only",
                            "undetermined"
                        )
                    )
                ),
            snv_indel = ifelse(str_detect(found_in, "INDEL"), "indel", "snv")
        )

    # Load epitope data
    tested_epitopes <- data.table("Mut Epitope" = readLines(tested_epitopes_files[[y]]))
    # Load epitope variant key
    epitope_variant_key <- fread(epitope_variant_key_files[[y]])
    # Annotate variant key if in jared's output or NA if not found
    tested_key <- tested_epitopes %>% left_join(epitope_variant_key %>% dplyr::select(`Mut Epitope`, `Variant key`), by = "Mut Epitope") %>%
        mutate(`Variant key` = ifelse(is.na(`Variant key`), "NA", `Variant key`))
    # Add variant key short for easier matching
    jared_key_epitope_tested <- tested_key %>% mutate(variant_key_short = gsub(" .*", "", `Variant key`))
    # # Homogenize variant key between Rebeca and Jared's outputs
    # jared_key_epitope_tested$`Variant key` <- gsub("-c.([ATCG-]+)\\d+([ATCG-])", " \\1>\\2", jared_key_epitope_tested$`Variant key`)
    # Split variant key into chrom and pos, and create a new variant_key_short, with position of deletions adjusted by -1 if Jared's notation ("ATCG>-") (to match VCF notation)
    jared_key_epitope_tested <- separate(jared_key_epitope_tested, variant_key_short, c("chrom", "pos"), ":") %>% separate(., `Variant key`, c("chrom_pos", "nt_change"), " ") %>%
            rowwise() %>%
            mutate(variant_key_short = ifelse(str_detect(`nt_change`, ">-"), paste0(chrom, ":", as.numeric(pos) - 1),
                paste0(chrom, ":", pos)))
    # readjust pos column
    jared_key_epitope_tested$pos <- strsplit(jared_key_epitope_tested$variant_key_short, ":") %>% sapply(., function(x) x[2])
    
    # Add info if variant has been tested by TMG
    ccf_tbl <- ccf_tbl %>% mutate(TMG_tested = ifelse(paste0(variant_key_short, nt_change) %in% paste0(jared_key_epitope_tested$variant_key_short, jared_key_epitope_tested$nt_change), "yes", "no"))

    # summary dna presence per tested variants
    joined <- jared_key_epitope_tested %>% left_join(ccf_tbl %>% dplyr::select(variant_key_short, dna_presence), by = "variant_key_short")
    joined$dna_presence[is.na(joined$dna_presence)] <- "NA"

    summary_tested_variants_dna_presence <- joined %>% distinct(chrom_pos, nt_change, .keep_all = TRUE) %>%
    group_by(dna_presence) %>% summarize(count = n()) %>%
        bind_rows(
            tibble(dna_presence = "Total_tested_variants", count = sum(.$count))
        ) %>% mutate(sample = ifelse(y == "Motricolor002", "GOIGROS005", y))

    # write table with categories of tested variants
    write.table(joined, file = paste0("tested_variants_dna_presence_categories_", y, ".tsv"), sep = "\t", quote = F, row.names = F)

    # return table		
    return(list(ccf_tbl = ccf_tbl, summary_tested = summary_tested_variants_dna_presence))
}
```

# Data table for all patients
```{r}
# Update samples to include GOIGROS005
samples <- c(
    "GOIGROS001",
    "GOIGROS002",
    "GOIGROS003",
    "GOIGROS004",
    "GOIGROS005",
    "GOIGROS008",
    "TIL012",
    "TIL015"
)

# Apply to all patients
list_ccf <- purrr::imap(files, get_ccf)

# Merge tables into one ccf table and keep only one row per mutation (not multiple rows if different epitopes)
ccf_table_pre <- list_ccf %>%  purrr::map("ccf_tbl") %>%
    bind_rows()

ccf_table <- ccf_table_pre %>%
    distinct(sample, POS, REF, ALT, immunogenic, .keep_all = TRUE)

# # Remove non-immunogenic LASP1 epitope row 1954  ***************** DOUBLE CHECK ROW IF CHANGING INPUTS!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
# ccf_table <- ccf_table[-1954, ]

# Remove "GROS" from sample names
ccf_table$sample <- gsub("GROS0", "-", ccf_table$sample)

# Change GOI-08 for GOI-06, TIL012 for GOI-07 and TIL015 for GOI-08
ccf_table$sample <- gsub("GOI-08", "GOI-06", ccf_table$sample)
ccf_table$sample <- gsub("TIL012", "GOI-07", ccf_table$sample)
ccf_table$sample <- gsub("TIL015", "GOI-08", ccf_table$sample)

# Export table
write.table(ccf_table, file = "GOI_cohort_variants_CCF.tsv", sep = "\t", quote = F, row.names = F)

```


# Mutation burden and coding consequences
### Number of Mutations
```{r}
ccf_summary <- ccf_table %>%
    group_by(sample, coding_consequence) %>%
    summarise(n_mutations = n()) %>%
    spread(key = coding_consequence, value = n_mutations, fill = 0) %>%
    mutate(total = non_synonymous + synonymous_or_noncoding)

datatable(ccf_summary)
```

```{r}
# Plot N
ggplot(ccf_summary %>% pivot_longer(
    cols = 2:3,
    names_to = "coding_consequence"
), aes(x = sample, y = value, fill = coding_consequence)) +
    geom_bar(stat = "identity", position = "stack") +
    labs(x = "Samples", y = "N Mutations", fill = "coding consequence") +
    paletteer::scale_fill_paletteer_d("nbapalettes::wizards_city") +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

### Percent of Mutations
```{r}
# Plot Percent
ccf_summary <- ccf_table %>%
    group_by(sample, coding_consequence) %>%
    summarise(n_mutations = n(), .groups = "drop") %>%
    group_by(sample) %>%
    mutate(percent = round(n_mutations / sum(n_mutations) * 100, 2)) %>%
    ungroup() %>%
    select(-n_mutations) %>%
    spread(key = coding_consequence, value = percent, fill = 0)

datatable(ccf_summary)

ggplot(ccf_summary %>% pivot_longer(
    cols = 2:3,
    names_to = "coding_consequence"
), aes(x = sample, y = value, fill = coding_consequence)) +
    geom_bar(stat = "identity", position = "stack") +
    labs(x = "Samples", y = "% Mutations", fill = "coding consequence") +
    paletteer::scale_fill_paletteer_d("nbapalettes::wizards_city") +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

# Shared vs. sample-specific mutations

## All mutations
### Number of Mutations
```{r}
ccf_summary <- ccf_table %>%
    group_by(sample, dna_presence) %>%
    summarise(n_mutations = n()) %>%
    spread(key = dna_presence, value = n_mutations, fill = 0)

datatable(ccf_summary)
```

```{r}
# Plot N
ggplot(ccf_summary %>% pivot_longer(
    cols = 2:4,
    names_to = "dna_presence"
), aes(x = sample, y = value, fill = dna_presence)) +
    geom_bar(stat = "identity", position = "stack") +
    labs(x = "Samples", y = "N", fill = "variant presence") +
    scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

### Percent of Mutations
```{r}
# Plot Percent
ccf_summary <- ccf_table %>%
    group_by(sample, dna_presence) %>%
    summarise(n_mutations = n(), .groups = "drop") %>%
    group_by(sample) %>%
    mutate(percent = round(n_mutations / sum(n_mutations) * 100, 2)) %>%
    ungroup() %>%
    # select(-n_mutations) %>%
    spread(key = dna_presence, value = percent, fill = 0)

#  All categories
ggplot(
    ccf_summary %>% pivot_longer(
        cols = 3:5,
        names_to = "dna_presence", values_to = "percent"
    ) %>%
        filter(percent > 0),
    aes(x = sample, y = percent, fill = dna_presence)
) +
    geom_bar(stat = "identity", position = "fill") +
    geom_text(aes(label = n_mutations), position = position_fill(vjust = 0.5), color = "white") +
    labs(x = "Samples", y = "% Mutations", fill = "variant presence") +
    scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
    ggtitle("Variant Presence") +
    scale_y_continuous(labels = function(x) round(x * 100, 1))

# Only Shared
ggplot(
    ccf_summary %>% pivot_longer(
        cols = 3:5,
        names_to = "dna_presence", values_to = "percent"
    ) %>%
        filter(percent > 0, dna_presence == "Shared"),
    aes(x = sample, y = percent)
) +
    geom_bar(stat = "identity", fill = "#7CC6BF") +
    geom_text(aes(label = n_mutations), position = position_stack(vjust = 0.5), color = "white") +
    labs(x = "Samples", y = "% Mutations", fill = "variant presence") +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
    theme(axis.title.x = element_blank()) +
    ggtitle("Percent of shared mutations (from all mutations)")


# Table Percent
ccf_summary <- ccf_table %>%
    group_by(sample, dna_presence) %>%
    summarise(n_mutations = n(), .groups = "drop") %>%
    group_by(sample) %>%
    mutate(percent = round(n_mutations / sum(n_mutations) * 100, 2)) %>%
    ungroup() %>%
    select(-n_mutations) %>%
    spread(key = dna_presence, value = percent, fill = 0)

datatable(ccf_summary)
```

## NSMs
### Number of Mutations
```{r}
ccf_summary <- ccf_table %>%
    filter(coding_consequence == "non_synonymous") %>%
    group_by(sample, dna_presence) %>%
    summarise(n_mutations = n()) %>%
    spread(key = dna_presence, value = n_mutations, fill = 0)

datatable(ccf_summary)
```

```{r}
# Plot N
ggplot(ccf_summary %>% pivot_longer(
    cols = 2:4,
    names_to = "dna_presence"
), aes(x = sample, y = value, fill = dna_presence)) +
    geom_bar(stat = "identity", position = "stack") +
    labs(x = "Samples", y = "N", fill = "variant presence") +
    scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

### Percent of Mutations
```{r}
# Plot Percent
ccf_summary <- ccf_table %>%
    filter(coding_consequence == "non_synonymous") %>%
    group_by(sample, dna_presence) %>%
    summarise(n_mutations = n(), .groups = "drop") %>%
    group_by(sample) %>%
    mutate(percent = round(n_mutations / sum(n_mutations) * 100, 2)) %>%
    ungroup() %>%
    # select(-n_mutations) %>%
    spread(key = dna_presence, value = percent, fill = 0)

#  All categories
ggplot(
    ccf_summary %>% pivot_longer(
        cols = 3:5,
        names_to = "dna_presence", values_to = "percent"
    ) %>%
        filter(percent > 0),
    aes(x = sample, y = percent, fill = dna_presence)
) +
    geom_bar(stat = "identity", position = "fill") +
    geom_text(aes(label = n_mutations), position = position_fill(vjust = 0.5), color = "white") +
    labs(x = "Samples", y = "% Mutations", fill = "variant presence") +
    scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
    ggtitle("Variant Presence") +
    scale_y_continuous(labels = function(x) round(x * 100, 1))

# Only Shared
ggplot(
    ccf_summary %>% pivot_longer(
        cols = 3:5,
        names_to = "dna_presence", values_to = "percent"
    ) %>%
        filter(percent > 0, dna_presence == "Shared"),
    aes(x = sample, y = percent)
) +
    geom_bar(stat = "identity", fill = "#7CC6BF") +
    geom_text(aes(label = n_mutations), position = position_stack(vjust = 0.5), color = "white") +
    labs(x = "Samples", y = "% Mutations", fill = "variant presence") +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
    theme(axis.title.x = element_blank()) +
    ggtitle("Percent of shared mutations (from NSM mutations)")

# Table percent
ccf_summary <- ccf_table %>%
    filter(coding_consequence == "non_synonymous") %>%
    group_by(sample, dna_presence) %>%
    summarise(n_mutations = n(), .groups = "drop") %>%
    group_by(sample) %>%
    mutate(percent = round(n_mutations / sum(n_mutations) * 100, 2)) %>%
    ungroup() %>%
    select(-n_mutations) %>%
    spread(key = dna_presence, value = percent, fill = 0)

datatable(ccf_summary)
```

```{r}
ggsave("Frequency_of_shared_NSM.pdf", height = 5, width = 4.5)
```

## Coding consequences
### Number of Mutations
```{r}
ccf_summary <- ccf_table %>%
    group_by(sample, dna_presence, coding_consequence) %>%
    summarise(n_mutations = n())

ccf_summary <- ccf_summary %>%
    tidyr::unite(col = "group", c("coding_consequence", "dna_presence"), sep = ".") %>%
    pivot_wider(
        names_from = group,
        values_from = n_mutations,
        values_fill = list(n_mutations = 0)
    )

datatable(ccf_summary)
```

### Percent of Mutations
```{r}
ccf_summary_percent <- ccf_table %>%
    group_by(sample, dna_presence, coding_consequence) %>%
    summarise(n_mutations = n(), .groups = "drop") %>%
    group_by(sample, coding_consequence) %>%
    mutate(percent = round(n_mutations / sum(n_mutations) * 100, 2)) %>%
    ungroup()


ggplot(
    ccf_summary_percent %>% filter(coding_consequence == "non_synonymous"),
    aes(x = dna_presence, y = n_mutations)
) +
    geom_bar(stat = "identity", aes(fill = dna_presence)) +
    geom_text(aes(label = paste0(percent, "%")), size = 2.5, vjust = -0.4) +
    scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
    facet_wrap(~sample, scales = "free") +
    scale_y_continuous(expand = expansion(mult = c(0.1, 0.5))) + # Otherwise we have the percentage geom_text hidden by the sample name boxes
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
    labs(fill = "Variant presence") +
    ylab("Number of NSM")

ggsave("Summary_Number_NSMs.pdf", width = 6, height = 6)


ccf_summary_percent <- ccf_summary_percent %>%
    select(-n_mutations) %>%
    tidyr::unite(col = "group", c("coding_consequence", "dna_presence"), sep = ".") %>%
    pivot_wider(
        names_from = group,
        values_from = percent,
        values_fill = list(percent = 0)
    )

datatable(ccf_summary_percent)
```

## SNV vs Indels
### Number of Mutations
```{r}
ccf_summary <- ccf_table %>%
    group_by(sample, dna_presence, snv_indel) %>%
    summarise(n_mutations = n())

ccf_summary <- ccf_summary %>%
    tidyr::unite(col = "group", c("dna_presence", "snv_indel"), sep = ".") %>%
    pivot_wider(
        names_from = group,
        values_from = n_mutations,
        values_fill = list(n_mutations = 0)
    )

datatable(ccf_summary)
```

# Lollipop plots
## All mutations
We are showing here the total number of mutations (all kind) for the FrTu (FrTu_only + Shared) and cfDNA (cfDNA + Shared) samples
```{r}
lollipop_table <- ccf_summary %>% mutate(total_FrTu = sum(c_across(matches("FrTu|Shared"))))
lollipop_table <- lollipop_table %>% mutate(total_cfDNA = sum(c_across(matches("cfDNA|Shared"))))

lollipop_Frtu_all <- ggplot(lollipop_table, aes(x = sample, y = log10(total_FrTu))) +
    geom_segment(aes(x = sample, xend = sample, y = 0, yend = log10(total_FrTu)), color = "#D9D8D4", linewidth = 1) +
    geom_point(color = "#D9D8D4", size = 5) +
    coord_flip() +
    theme_classic() +
    theme(
        plot.title = element_blank(),
        axis.text = element_text(size = 12),
        legend.position = "none",
        axis.text.y = element_text(margin = margin(0, 10, 0, 0), hjust = 0, vjust = 0.5),
    ) +
    geom_text(aes(label = total_FrTu), size = 3.5, vjust = -1.25) +
    ylim(0, 5) +
    theme(aspect.ratio = 1) +
    theme(
        axis.title.y = element_blank()
    )

lollipop_Frtu_all <- lollipop_Frtu_all + labs(y = "FrTu Total Mutations (Log10)")

# Lollipop cfDNA_all_mutations
lollipop_cfDNA_all <- ggplot(lollipop_table, aes(x = sample, y = log10(total_cfDNA))) +
    geom_segment(aes(x = sample, xend = sample, y = 0, yend = log10(total_cfDNA)), color = "#FCC557", linewidth = 1) +
    geom_point(color = "#FCC557", size = 5) +
    coord_flip() +
    theme_classic() +
    theme(
        plot.title = element_blank(),
        axis.text = element_text(size = 12),
        legend.position = "none"
    ) +
    geom_text(aes(label = total_cfDNA), size = 3.5, vjust = -1.25) +
    scale_y_reverse(lim = c(5, 0)) +
    theme(aspect.ratio = 1) +
    theme(
        axis.title.y = element_blank(),
        axis.text.y = element_blank()
    ) +
    scale_x_discrete(position = "top")

lollipop_cfDNA_all <- lollipop_cfDNA_all + labs(y = "cfDNA Total Mutations (Log10)")

patchwork <- lollipop_cfDNA_all + lollipop_Frtu_all

patchwork
```

```{r}
# Save lollipop as pdf
ggsave("lollipop_all_mut.pdf")
```

## NSMs
We are showing here the total number of NSMs for the FrTu (FrTu_only + Shared) and cfDNA (cfDNA + Shared) samples

```{r}
ccf_summary <- ccf_table %>%
    group_by(sample, dna_presence, coding_consequence) %>%
    summarise(n_mutations = n())

ccf_summary <- ccf_summary %>%
    tidyr::unite(col = "group", c("coding_consequence", "dna_presence"), sep = ".") %>%
    pivot_wider(
        names_from = group,
        values_from = n_mutations,
        values_fill = list(n_mutations = 0)
    )
```

```{r}
lollipop_table <- ccf_summary %>% mutate(nsm_FrTu = sum(c_across(matches("non_synonymous.FrTu|non_synonymous.Shared"))))
lollipop_table <- lollipop_table %>% mutate(nsm_cfDNA = sum(c_across(matches("non_synonymous.cfDNA|non_synonymous.Shared"))))

lollipop_Frtu_all <- ggplot(lollipop_table, aes(x = sample, y = log10(nsm_FrTu))) +
    geom_segment(aes(x = sample, xend = sample, y = 0, yend = log10(nsm_FrTu)), color = "#D9D8D4", linewidth = 1) +
    geom_point(color = "#D9D8D4", size = 5) +
    coord_flip() +
    theme_classic() +
    theme(
        plot.title = element_blank(),
        axis.text = element_text(size = 12),
        legend.position = "none",
        axis.text.y = element_text(margin = margin(0, 10, 0, 0), hjust = 0, vjust = 0.5),
    ) +
    geom_text(aes(label = nsm_FrTu), size = 3.5, vjust = -1.25) +
    ylim(0, 5) +
    theme(aspect.ratio = 1) +
    theme(
        axis.title.y = element_blank()
    )

lollipop_Frtu_all <- lollipop_Frtu_all + labs(y = "FrTu NSM Mutations (Log10)")

# Lollipop cfDNA_all_mutations
lollipop_cfDNA_all <- ggplot(lollipop_table, aes(x = sample, y = log10(nsm_cfDNA))) +
    geom_segment(aes(x = sample, xend = sample, y = 0, yend = log10(nsm_cfDNA)), color = "#FCC557", linewidth = 1) +
    geom_point(color = "#FCC557", size = 5) +
    coord_flip() +
    theme_classic() +
    theme(
        plot.title = element_blank(),
        axis.text = element_text(size = 12),
        legend.position = "none"
    ) +
    geom_text(aes(label = nsm_cfDNA), size = 3.5, vjust = -1.25) +
    scale_y_reverse(lim = c(5, 0)) +
    theme(aspect.ratio = 1) +
    theme(
        axis.title.y = element_blank(),
        axis.text.y = element_blank()
    ) +
    scale_x_discrete(position = "top")

lollipop_cfDNA_all <- lollipop_cfDNA_all + labs(y = "cfDNA NSM Mutations (Log10)")

patchwork <- lollipop_cfDNA_all + lollipop_Frtu_all

patchwork
```

```{r}
# Save lollipop as pdf
ggsave("lollipop_NSM.pdf")
```


# Categories of TMG tested variants
```{r}
# Merge list tables into one summary table 
summary_tested_table <- list_ccf %>%  purrr::map("summary_tested") %>%
    bind_rows()

# Remove "GROS" from sample names
summary_tested_table$sample <- gsub("GROS0", "-", summary_tested_table$sample)

# Change GOI-08 for GOI-06, TIL012 for GOI-07 and TIL015 for GOI-08
summary_tested_table$sample <- gsub("GOI-08", "GOI-06", summary_tested_table$sample)
summary_tested_table$sample <- gsub("TIL012", "GOI-07", summary_tested_table$sample)
summary_tested_table$sample <- gsub("TIL015", "GOI-08", summary_tested_table$sample)

datatable(summary_tested_table)
```

```{r}
# Plots
percent_table <- summary_tested_table %>% filter(dna_presence != "Total_tested_variants") %>%
  group_by(sample) %>%
  mutate(
    percent = round(100 * count / sum(count), 2)
  ) %>%
  ungroup()

ggplot(
    percent_table,
    aes(x = sample, y = percent, fill = dna_presence)
) +
    geom_bar(stat = "identity", position = "fill") +
    geom_text(aes(label = count), position = position_fill(vjust = 0.5), color = "white") +
    labs(x = "Samples", y = "% Mutations", fill = "variant presence") +
    scale_fill_manual(values = c("#FCC557", "#D9D8D4", "grey40", "#7CC6BF")) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
    ggtitle("Variant Presence") +
    scale_y_continuous(labels = function(x) round(x * 100, 1))

ggsave("Summary_TMG_tested_variants.pdf", width = 7, height = 6)
```

# CCF plots

```{r, echo = TRUE}
# Filter ccf_table to keep mutations with min_cov in all three samples (FrTu, cfDNA and Normal) and filter out mutations with CN == 0 or NA
ccf_table <- ccf_table %>% 
    dplyr::filter(
        cov_FrTu >= min_cov & cov_cfDNA >= min_cov & cov_Normal >= min_cov & nMajor_cfDNA != 0 & nMajor_FrTu != 0
    )
```

### No limit at CCF=1
```{r, fig.width=10, fig.height=6.5}
p <- ggplot(ccf_table, aes(x = ccf_FrTu, y = ccf_cfDNA, color = dna_presence)) +
    geom_point(alpha = 0.3) +
    facet_wrap(~sample, scales = "free") +
    scale_color_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
    theme_classic()

p
```

<br>
<br>

### CCF plots with density

```{r}
# Add gene name info in TIL012/GOI-07 and TIL015/GOI-08 samples (because ran in our pipeline and not in Jared's like GOI patients)
ccf_table$`gene_name` <- ifelse(
    ccf_table$sample %in% c("GOI-07", "GOI-08"), gsub("missense_variant.*_([A-Z0-9]*)_ENST.*", "\\1", ccf_table$`Effects`),
    ifelse(
        ccf_table$sample == "GOI-06", gsub(".*_([A-Z0-9]+)_uc.*", "\\1", ccf_table$`Effects`),
        ccf_table$`Gene Name`))
```

```{r ccf plots with density no limit at 1}
samples <- unique(ccf_table$sample) # necessary as we changed the names!
for (i in samples) {
    plot_table <- ccf_table %>%
        filter(sample == i)

    p <- ggplot(
        plot_table,
        aes(
            x = ccf_FrTu, y = ccf_cfDNA, color = dna_presence,
            text = paste0(
                "REF_counts_FrTu: ", REF_counts_FrTu, "\n",
                "ALT_counts_FrTu: ", ALT_counts_FrTu, "\n",
                "REF_counts_cfDNA: ", REF_counts_cfDNA, "\n",
                "ALT_counts_cfDNA: ", ALT_counts_cfDNA, "\n",
                "REF_counts_Normal: ", REF_counts_Normal, "\n",
                "ALT_counts_Normal: ", ALT_counts_Normal, "\n",
                "nMajor_FrTu: ", nMajor_FrTu, "\n",
                "nMinor_FrTu: ", nMinor_FrTu, "\n",
                "nMajor_cfDNA: ", nMajor_cfDNA, "\n",
                "nMinor_cfDNA: ", nMinor_cfDNA, "\n"
            )
        )
    ) +
        geom_point(alpha = 0.3) +
        scale_color_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
        theme_classic() +
        theme(legend.position = "none")

    dens1 <- ggplot(plot_table, aes(x = ccf_FrTu, y = dna_presence, fill = dna_presence)) +
        geom_density_ridges(alpha = 0.4) +
        scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
        theme_void() +
        theme(legend.position = "none")

    dens2 <- ggplot(plot_table, aes(x = ccf_cfDNA, y = dna_presence, fill = dna_presence)) +
        geom_density_ridges(alpha = 0.4) +
        theme(legend.position = "none") +
        coord_flip() +
        scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
        theme_void()

    hist1 <- ggplot(plot_table, aes(x = ccf_FrTu, fill = dna_presence)) +
        geom_histogram(binwidth = 0.05) +
        scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
        theme_classic() +
        theme(
            axis.title.x = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.line.x = element_blank()
        ) +
        scale_y_reverse() +
        theme(legend.position = "none")

    hist2 <- ggplot(plot_table, aes(x = ccf_cfDNA, fill = dna_presence)) +
        geom_histogram(binwidth = 0.05) +
        scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
        theme_classic() +
        theme(
            axis.title.y = element_blank(),
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.line.y = element_blank(),
            axis.text.x = element_text(angle = 90)
        ) +
        coord_flip() +
        scale_y_reverse() +
        theme(legend.position = "none")

    patch_plot <- plot_spacer() + dens1 + plot_spacer() +
        hist2 + p + dens2 +
        plot_spacer() + hist1 + plot_spacer() +
        plot_layout(ncol = 3, nrow = 3, widths = c(1, 5, 1), heights = c(1, 5, 1)) +
        plot_annotation(title = i)

    ggsave(plot = patch_plot, filename = paste0("CCF_density_", i, ".png"))

    # # 	# export plotly
    # widget <- ggplotly(p)
    # htmlwidgets::saveWidget(widget, file = paste0("CCF_density_", i, ".html"))
}
```


```{r ccf plots with density limit at 1}
for (i in samples) {
    plot_table <- ccf_table %>%
        mutate(
            ccf_FrTu = ifelse(ccf_FrTu >= 1, 1, ccf_FrTu),
            ccf_cfDNA = ifelse(ccf_cfDNA >= 1, 1, ccf_cfDNA)
        ) %>%
        filter(sample == i)

    p <- ggplot(
        plot_table,
        aes(x = ccf_FrTu, y = ccf_cfDNA, color = dna_presence)
    ) +
        geom_point(alpha = 0.3) +
        scale_color_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
        geom_point(plot_table %>% dplyr::filter(immunogenic == "yes"), mapping = aes(x = ccf_FrTu, y = ccf_cfDNA), shape = 8, color = "#333333", size = 2) +
        geom_text_repel(plot_table %>% dplyr::filter(immunogenic == "yes"), mapping = aes(x = ccf_FrTu, y = ccf_cfDNA, label = gene_name),
        color = "grey30", size = 3.3,
        min.segment.length = unit(0, "lines"), segment.color = "grey",
        max.overlaps = Inf,
        box.padding = 0.5,
        fontface = "italic", force = 100, max.time = 50
    ) +
        theme_classic() +
        scale_x_continuous(
        limits = c(-0.1, 1.2),
        breaks = c(0, 0.25, 0.5, 0.75, 1)
    ) +
    scale_y_continuous(
        limits = c(-0.1, 1.2),
        breaks = c(0, 0.25, 0.5, 0.75, 1)
    ) +
        theme(legend.position = "none")

    dens1 <- ggplot(plot_table, aes(x = ccf_FrTu, y = dna_presence, fill = dna_presence)) +
        geom_density_ridges(alpha = 0.4) +
        xlim(-0.1, 1.2) +
        scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
        theme_void() +
        theme(legend.position = "none")

    dens2 <- ggplot(plot_table, aes(x = ccf_cfDNA, y = dna_presence, fill = dna_presence)) +
        geom_density_ridges(alpha = 0.4) +
        theme(legend.position = "none") +
        xlim(-0.1, 1.2) +
        coord_flip() +
        scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
        theme_void()

    hist1 <- ggplot(plot_table, aes(x = ccf_FrTu, fill = dna_presence)) +
        geom_histogram(binwidth = 0.05) +
        xlim(-0.1, 1.2) +
        scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
        theme_classic() +
        theme(
            axis.title.x = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.line.x = element_blank()
        ) +
        scale_y_reverse() +
        theme(legend.position = "none")

    hist2 <- ggplot(plot_table, aes(x = ccf_cfDNA, fill = dna_presence)) +
        geom_histogram(binwidth = 0.05) +
        xlim(-0.1, 1.2) +
        scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
        theme_classic() +
        theme(
            axis.title.y = element_blank(),
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.line.y = element_blank(),
            axis.text.x = element_text(angle = 90)
        ) +
        coord_flip() +
        scale_y_reverse() +
        theme(legend.position = "none")

    patch_plot <- plot_spacer() + dens1 + plot_spacer() +
        hist2 + p + dens2 +
        plot_spacer() + hist1 + plot_spacer() +
        plot_layout(ncol = 3, nrow = 3, widths = c(1, 5, 1), heights = c(1, 5, 1)) +
        plot_annotation(title = paste0(i, " CCF max == 1"))

    ggsave(plot = patch_plot, filename = paste0("CCF_density_", i, "_CCFmax1.png"))
    ggsave(plot = patch_plot, filename = paste0("CCF_density_", i, "_CCFmax1.pdf"))
}
```

```{r out.width = "100%", fig.align='center', fig.width=10, fig.height=9}
# Include plot in R Markdown html report
files <- list.files(
    path = getwd(),
    pattern = "^CCF_density_(.*).png$",
    full.names = TRUE
)
knitr::include_graphics(files, error = FALSE)
```

# VAF plots
```{r vaf plots with density}
for (i in samples) {
    plot_table <- ccf_table %>%
        # mutate(
        #     vaf_FrTu = vaf_FrTu * 100,
        #     vaf_cfDNA = vaf_cfDNA * 100
        # ) %>%
        filter(sample == i)
    
    # Change level to have shared in the background  
    plot_table$dna_presence <- factor(plot_table$dna_presence, levels = c("Shared", "cfDNA_only", "FrTu_only"))
    plot_table <- plot_table %>% arrange(dna_presence)
    
    p <- ggplot(
        plot_table,
        aes(x = vaf_FrTu, y = vaf_cfDNA, color = dna_presence)
    ) +
        geom_point(alpha = 0.3) +
        scale_color_manual(values = c("#7CC6BF", "#FCC557", "#D9D8D4")) +
        geom_point(plot_table %>% dplyr::filter(immunogenic == "yes"), mapping = aes(x = vaf_FrTu, y = vaf_cfDNA), shape = 8, color = "#333333", size = 2) +
    #     geom_text_repel(plot_table %>% dplyr::filter(immunogenic == "yes"), mapping = aes(x = vaf_FrTu, y = vaf_cfDNA, label = `Gene Name`),
    #     color = "grey30", size = 3.3,
    #     min.segment.length = unit(0, "lines"), segment.color = "grey",
    #     max.overlaps = Inf,
    #     box.padding = 0.5,
    #     fontface = "italic", force = 100, max.time = 50
    # ) +
        theme_classic() +
        scale_x_continuous(
        limits = c(-0.1, 1.2),
        breaks = c(0, 0.25, 0.5, 0.75, 1)
    ) +
    scale_y_continuous(
        limits = c(-0.1, 1.2),
        breaks = c(0, 0.25, 0.5, 0.75, 1)
    ) +
        theme(legend.position = "none")

    dens1 <- ggplot(plot_table, aes(x = vaf_FrTu, y = dna_presence, fill = dna_presence)) +
        geom_density_ridges(alpha = 0.4) +
        xlim(-0.1, 1.2) +
        scale_fill_manual(values = c("#7CC6BF", "#FCC557", "#D9D8D4")) +
        theme_void() +
        theme(legend.position = "none")

    dens2 <- ggplot(plot_table, aes(x = vaf_cfDNA, y = dna_presence, fill = dna_presence)) +
        geom_density_ridges(alpha = 0.4) +
        theme(legend.position = "none") +
        xlim(-0.1, 1.2) +
        coord_flip() +
        scale_fill_manual(values = c("#7CC6BF", "#FCC557", "#D9D8D4")) +
        theme_void()

    hist1 <- ggplot(plot_table, aes(x = vaf_FrTu, fill = dna_presence)) +
        geom_histogram(binwidth = 0.05) +
        xlim(-0.1, 1.2) +
        scale_fill_manual(values = c("#7CC6BF", "#FCC557", "#D9D8D4")) +
        theme_classic() +
        theme(
            axis.title.x = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.line.x = element_blank()
        ) +
        scale_y_reverse() +
        theme(legend.position = "none")

    hist2 <- ggplot(plot_table, aes(x = vaf_cfDNA, fill = dna_presence)) +
        geom_histogram(binwidth = 0.05) +
        xlim(-0.1, 1.2) +
        scale_fill_manual(values = c("#7CC6BF", "#FCC557", "#D9D8D4")) +
        theme_classic() +
        theme(
            axis.title.y = element_blank(),
            axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            axis.line.y = element_blank(),
            axis.text.x = element_text(angle = 90)
        ) +
        coord_flip() +
        scale_y_reverse() +
        theme(legend.position = "none")

    patch_plot <- plot_spacer() + dens1 + plot_spacer() +
        hist2 + p + dens2 +
        plot_spacer() + hist1 + plot_spacer() +
        plot_layout(ncol = 3, nrow = 3, widths = c(1, 5, 1), heights = c(1, 5, 1)) +
        plot_annotation(title = paste0(i, " VAF"))

    ggsave(plot = patch_plot, filename = paste0("VAF_density_", i, ".png"))
    ggsave(plot = patch_plot, filename = paste0("VAF_density_", i, ".pdf"))
}
```

```{r}
# VAF plots GOI-05 ZOOMED
plot_table <- ccf_table %>%
        # mutate(
        #     vaf_FrTu = vaf_FrTu * 100,
        #     vaf_cfDNA = vaf_cfDNA * 100
        # ) %>%
        filter(sample == "GOI-05")

# Change level to have shared in the background  
plot_table$dna_presence <- factor(plot_table$dna_presence, levels = c("Shared", "cfDNA_only", "FrTu_only"))
plot_table <- plot_table %>% arrange(dna_presence)

p <- ggplot(
    plot_table,
    aes(x = vaf_FrTu, y = vaf_cfDNA, color = dna_presence)
) +
    geom_point(alpha = 0.3) +
    scale_color_manual(values = c("#7CC6BF", "#FCC557", "#D9D8D4")) +
    # geom_point(plot_table %>% dplyr::filter(immunogenic == "yes"), mapping = aes(x = vaf_FrTu, y = vaf_cfDNA), shape = 8, color = "#333333", size = 2) +
#     geom_text_repel(plot_table %>% dplyr::filter(immunogenic == "yes"), mapping = aes(x = vaf_FrTu, y = vaf_cfDNA, label = `Gene Name`),
#     color = "grey30", size = 3.3,
#     min.segment.length = unit(0, "lines"), segment.color = "grey",
#     max.overlaps = Inf,
#     box.padding = 0.5,
#     fontface = "italic", force = 100, max.time = 50
# ) +
    theme_classic() +
    scale_x_continuous(
    limits = c(-0.05, 0.55),
    breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5)
) +
scale_y_continuous(
    limits = c(-0.05, 0.55),
    breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5)
) +
    theme(legend.position = "none")

dens1 <- ggplot(plot_table, aes(x = vaf_FrTu, y = dna_presence, fill = dna_presence)) +
    geom_density_ridges(alpha = 0.4) +
    xlim(-0.05, 0.55) +
    scale_fill_manual(values = c("#7CC6BF", "#FCC557", "#D9D8D4")) +
    theme_void() +
    theme(legend.position = "none")

dens2 <- ggplot(plot_table, aes(x = vaf_cfDNA, y = dna_presence, fill = dna_presence)) +
    geom_density_ridges(alpha = 0.4) +
    theme(legend.position = "none") +
    xlim(-0.05, 0.55) +
    coord_flip() +
    scale_fill_manual(values = c("#7CC6BF", "#FCC557", "#D9D8D4")) +
    theme_void()

hist1 <- ggplot(plot_table, aes(x = vaf_FrTu, fill = dna_presence)) +
    geom_histogram(binwidth = 0.05) +
    xlim(-0.05, 0.55) +
    scale_fill_manual(values = c("#7CC6BF", "#FCC557", "#D9D8D4")) +
    theme_classic() +
    theme(
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank()
    ) +
    scale_y_reverse() +
    theme(legend.position = "none")

hist2 <- ggplot(plot_table, aes(x = vaf_cfDNA, fill = dna_presence)) +
    geom_histogram(binwidth = 0.05) +
    xlim(-0.05, 0.55) +
    scale_fill_manual(values = c("#7CC6BF", "#FCC557", "#D9D8D4")) +
    theme_classic() +
    theme(
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        axis.text.x = element_text(angle = 90)
    ) +
    coord_flip() +
    scale_y_reverse() +
    theme(legend.position = "none")

patch_plot <- plot_spacer() + dens1 + plot_spacer() +
    hist2 + p + dens2 +
    plot_spacer() + hist1 + plot_spacer() +
    plot_layout(ncol = 3, nrow = 3, widths = c(1, 5, 1), heights = c(1, 5, 1)) +
    plot_annotation(title = paste0("GOI-05", " VAF Zoomed"))

ggsave(plot = patch_plot, filename = paste0("VAF_density_GOI-05_ZOOM.png"))
ggsave(plot = patch_plot, filename = paste0("VAF_density_GOI-05_ZOOM.pdf"))
```

```{r out.width = "100%", fig.align='center', fig.width=10, fig.height=9}
# Include plot in R Markdown html report
files <- list.files(
    path = getwd(),
    pattern = "^VAF_density_(.*).png$",
    full.names = TRUE
)
knitr::include_graphics(files, error = FALSE)
```

# DNA presence Shared according to clonality
#### In FrTu
```{r}
plot_table <- ccf_table %>%
    filter(coding_consequence == "non_synonymous", dna_presence %in% c("Shared", "FrTu_only")) %>%
    mutate(
        FrTu_clonal = ifelse(ccf_FrTu >= 0.9, "clonal", "subclonal")
    )

detected_cfDNA <- ccf_table %>% filter(coding_consequence == "non_synonymous", dna_presence %in% c("Shared", "cfDNA_only"))
n_detected_cfDNA <- detected_cfDNA %>% group_by(sample) %>% summarise(n_detected_cfDNA = n())

# Plot Percent
plot_summary <- plot_table %>%
    group_by(sample, FrTu_clonal, dna_presence) %>%
    summarise(n_mutations = n(), .groups = "drop") %>%
    complete(sample, FrTu_clonal, dna_presence, fill = list(n_mutations = 0)) %>% # ensures entry for GOI-04 FrTu clonal even if 0 mutations
    group_by(sample, FrTu_clonal) %>%
    mutate(percent = round(n_mutations / sum(n_mutations) * 100, 2)) %>%
    ungroup()
# Change labels dna_presence to be more easy to understand
plot_summary$dna_presence <- gsub("FrTu_only", "Not detected in cfDNA", plot_summary$dna_presence)
plot_summary$dna_presence <- gsub("Shared", "Detected in cfDNA", plot_summary$dna_presence)

# Merge to add N Mutations detected in cfDNA samples
plot_summary <- plot_summary %>% left_join(n_detected_cfDNA, by = "sample")

# Plot with GOI-02 and GOI-03
ggplot(plot_summary, aes(x = FrTu_clonal, y = percent)) +
    geom_line(aes(group = sample), linewidth = 0.7, alpha = 0.4) +
    geom_point(aes(color = sample, size = log10(n_detected_cfDNA)), alpha = 0.7) +
    scale_color_manual(values = c("#00B8AAFF", "#000000", "#FD625EFF", "#F2C811FF", "#5F6B6DFF", "#8AD4EBFF", "#FE9666FF", "#A66999FF")) +
    facet_wrap(~dna_presence) +
    labs(
      x = "Clonality in Tumor Biopsy", 
      y = "% NSMs", 
      color = "Patient", 
      size = "Total NSMs in cfDNA"
    ) +
    scale_size_continuous(
      breaks = log10(c(10, 50, 250, 1000, 5000)),  # Log-transformed values
      labels = c(10, 50, 250, 1000, 5000)  # Linear values for legend
    ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    theme_classic()

ggsave("cfDNA_detection_by_FrTu_clonality.pdf", width = 6, height = 4)

# Function to run paired t-test
run_paired_ttest <- function(plot_summary) {
  plot_summary_wide <- plot_summary %>%
    select(sample, FrTu_clonal, percent) %>%
    pivot_wider(names_from = FrTu_clonal, values_from = percent)  # Wide format for paired test
  
  # Check if both "clonal" and "subclonal" exist in this subset
  if (all(c("clonal", "subclonal") %in% names(plot_summary_wide))) {
    ttest <- t.test(plot_summary_wide$clonal, plot_summary_wide$subclonal, paired = TRUE)
    return(data.frame(dna_presence = unique(plot_summary$dna_presence), p_value = ttest$p.value))
  } else {
    return(data.frame(dna_presence = unique(plot_summary$dna_presence), p_value = NA))
  }
}

# Apply test for each dna_presence group
paired_ttest_results <- plot_summary %>%
  group_by(dna_presence) %>%
  group_split() %>%
  lapply(run_paired_ttest) %>%
  bind_rows()

print(paired_ttest_results)

# Plot without GOI-02 and GOI-03
plot_summary <- plot_summary %>% filter(!sample %in% c("GOI-02", "GOI-03"))

ggplot(plot_summary, aes(x = FrTu_clonal, y = percent, color = sample)) +
	paletteer::scale_color_paletteer_d("Redmonder::qPBI") +
    geom_point(size = 3) +
    geom_line(aes(group = sample), linewidth = 1, alpha = 0.5) +
    facet_wrap(~dna_presence) +
    labs(x = "Clonality in Tumor Biopsy", y = "% Detected Mutations", color = "Patient") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    theme_classic()

ggsave("cfDNA_detection_by_FrTu_clonality_exclude_GOI-02-03.pdf", width = 5, height = 4)

# Function to run paired t-test
run_paired_ttest <- function(plot_summary) {
  plot_summary_wide <- plot_summary %>%
    select(sample, FrTu_clonal, percent) %>%
    pivot_wider(names_from = FrTu_clonal, values_from = percent)  # Wide format for paired test
  
  # Check if both "clonal" and "subclonal" exist in this subset
  if (all(c("clonal", "subclonal") %in% names(plot_summary_wide))) {
    ttest <- t.test(plot_summary_wide$clonal, plot_summary_wide$subclonal, paired = TRUE)
    return(data.frame(dna_presence = unique(plot_summary$dna_presence), p_value = ttest$p.value))
  } else {
    return(data.frame(dna_presence = unique(plot_summary$dna_presence), p_value = NA))
  }
}

# Apply test for each dna_presence group
paired_ttest_results <- plot_summary %>%
  group_by(dna_presence) %>%
  group_split() %>%
  lapply(run_paired_ttest) %>%
  bind_rows()

print(paired_ttest_results)

```

#### In cfDNA
```{r}
plot_table <- ccf_table %>%
    filter(coding_consequence == "non_synonymous", dna_presence %in% c("Shared", "cfDNA_only")) %>%
    mutate(
        cfDNA_clonal = ifelse(ccf_cfDNA >= 0.9, "clonal", "subclonal")
    )

detected_FrTu <- ccf_table %>% filter(coding_consequence == "non_synonymous", dna_presence %in% c("Shared", "FrTu_only"))
n_detected_FrTu <- detected_FrTu %>% group_by(sample) %>% summarise(n_detected_FrTu = n())

# Plot Percent
plot_summary <- plot_table %>%
    group_by(sample, cfDNA_clonal, dna_presence) %>%
    summarise(n_mutations = n(), .groups = "drop") %>%
    complete(sample, cfDNA_clonal, dna_presence, fill = list(n_mutations = 0)) %>% # ensures entry for GOI-04 cfDNA clonal even if 0 mutations
    group_by(sample, cfDNA_clonal) %>%
    mutate(percent = round(n_mutations / sum(n_mutations) * 100, 2)) %>%
    ungroup()
plot_summary <- plot_summary %>% filter(!sample %in% c("GOI-02", "GOI-03"))
# Change labels dna_presence to be more easy to understand
plot_summary$dna_presence <- gsub("cfDNA_only", "Not detected in FrTu", plot_summary$dna_presence)
plot_summary$dna_presence <- gsub("Shared", "Detected in FrTu", plot_summary$dna_presence)

# Merge to add N Mutations detected in FrTu samples
plot_summary <- plot_summary %>% left_join(n_detected_FrTu, by = "sample")

# Plot without GOI-02 and GOI-03 (as no ctDNA in cfDNA, doesn't make sense to check at clonality in cfDNA)
ggplot(plot_summary, aes(x = cfDNA_clonal, y = percent)) +
    geom_line(aes(group = sample), linewidth = 0.7, alpha = 0.4) +
    geom_point(aes(color = sample, size = log10(n_detected_FrTu)), alpha = 0.7) +
    scale_color_manual(values = c("#00B8AAFF", "#000000", "#FD625EFF", "#F2C811FF", "#5F6B6DFF", "#8AD4EBFF", "#FE9666FF", "#A66999FF")) +
    facet_wrap(~dna_presence) +
    labs(
      x = "Clonality in Liquid Biopsy", 
      y = "% NSMs", 
      color = "Patient", 
      size = "Total NSMs in FrTu"
    ) +
    scale_size_continuous(
      breaks = log10(c(10, 50, 250, 1000, 5000)),  # Log-transformed values
      labels = c(10, 50, 250, 1000, 5000)  # Linear values for legend
    ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    theme_classic()

ggsave("FrTu_detection_by_cfDNA_clonality.pdf", width = 6, height = 4)

# Function to run paired t-test
run_paired_ttest <- function(plot_summary) {
  plot_summary_wide <- plot_summary %>%
    select(sample, cfDNA_clonal, percent) %>%
    pivot_wider(names_from = cfDNA_clonal, values_from = percent)  # Wide format for paired test
  
  # Check if both "clonal" and "subclonal" exist in this subset
  if (all(c("clonal", "subclonal") %in% names(plot_summary_wide))) {
    ttest <- t.test(plot_summary_wide$clonal, plot_summary_wide$subclonal, paired = TRUE)
    return(data.frame(dna_presence = unique(plot_summary$dna_presence), p_value = ttest$p.value))
  } else {
    return(data.frame(dna_presence = unique(plot_summary$dna_presence), p_value = NA))
  }
}

# Apply test for each dna_presence group
paired_ttest_results <- plot_summary %>%
  group_by(dna_presence) %>%
  group_split() %>%
  lapply(run_paired_ttest) %>%
  bind_rows()

print(paired_ttest_results)

```

# Summaries Immunogenic Mutations
```{r}
# # Regenerate ccf_table with all mutations and all transcripts (otherwise we might lose an immunogenic mutation if other transcript affected, other epitope created not immunogenic)
# ccf_table <- list_ccf %>%
#     bind_rows()



# Plot
p <- ggplot(
    ccf_table %>%
        mutate(
            ccf_FrTu = ifelse(ccf_FrTu >= 1, 1, ccf_FrTu),
            ccf_cfDNA = ifelse(ccf_cfDNA >= 1, 1, ccf_cfDNA)
        ) %>%
        filter(immunogenic == "yes") %>% distinct(sample, gene_name, .keep_all = TRUE),
    aes(x = ccf_FrTu, y = ccf_cfDNA, color = dna_presence) #, shape = sample
) +
    geom_beeswarm(size = 3) +
    scale_color_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
    geom_text_repel(aes(label = gene_name),
        color = "grey30", size = 3.3,
        min.segment.length = unit(0, "lines"), segment.color = "grey",
        #   nudge_x = -15000,
        #   hjust = 0,
        #   direction = "y",
        max.overlaps = Inf,
        box.padding = 0.5,
        fontface = "italic", force = 10,
    ) +
    theme_classic() +
    labs(color = "Variant presence", shape = "Patient")

p
# ggplotly(p) %>% layout(width = 1000, height = 600)

ggsave("Immunogenic_CCF.pdf", width = 7.6, height = 6)

# Plot with patient name
p <- ggplot(
    ccf_table %>%
        mutate(
            ccf_FrTu = ifelse(ccf_FrTu >= 1, 1, ccf_FrTu),
            ccf_cfDNA = ifelse(ccf_cfDNA >= 1, 1, ccf_cfDNA)
        ) %>%
        filter(immunogenic == "yes") %>% distinct(sample, gene_name, .keep_all = TRUE),
    aes(x = ccf_FrTu, y = ccf_cfDNA, fill = dna_presence)
) +
    geom_point(size = 3, shape = 21, color = "white") +
    scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
    geom_text_repel(aes(label = gene_name, color = sample),
        size = 3.3,
        min.segment.length = unit(0, "lines"), segment.color = "grey",
        #   nudge_x = -15000,
        #   hjust = 0,
        #   direction = "y",
        max.overlaps = Inf,
        box.padding = 0.5,
        fontface = "italic", force = 30
    ) +
    paletteer::scale_color_paletteer_d("rcartocolor::Vivid") +
    # paletteer::scale_color_paletteer_d("beyonce::X18") +
    theme_classic() +
    labs(fill = "Variant presence", color = "Patient")

p
# ggplotly(p) %>% layout(width = 1000, height = 600)

ggsave("Immunogenic_CCF_with_patient_name.pdf", width = 7.6, height = 6)


# Plot uncapped
p <- ggplot(
    ccf_table %>%
        # mutate(
        #     ccf_FrTu = ifelse(ccf_FrTu >= 1, 1, ccf_FrTu),
        #     ccf_cfDNA = ifelse(ccf_cfDNA >= 1, 1, ccf_cfDNA)
        # ) %>%
        filter(immunogenic == "yes") %>% distinct(sample, gene_name, .keep_all = TRUE),
    aes(x = ccf_FrTu, y = ccf_cfDNA, color = dna_presence) #, shape = sample
) +
    geom_beeswarm(size = 3) +
    scale_color_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
    geom_text_repel(aes(label = gene_name),
        color = "grey30", size = 3.3,
        min.segment.length = unit(0, "lines"), segment.color = "grey",
        #   nudge_x = -15000,
        #   hjust = 0,
        #   direction = "y",
        max.overlaps = Inf,
        box.padding = 0.5,
        fontface = "italic", force = 100, max.time = 50
    ) +
    theme_classic() +
    labs(color = "Variant presence", shape = "Patient", x = "Mutation copy number (Tumor Biopsy)", y = "Mutation copy number (cfDNA)")

p
# ggplotly(p) %>% layout(width = 1000, height = 600)

ggsave("Immunogenic_CCF_uncapped.pdf", width = 7.6, height = 6)


# Plot uncapped with patient name
p <- ggplot(
    ccf_table %>%
        # mutate(
        #     ccf_FrTu = ifelse(ccf_FrTu >= 1, 1, ccf_FrTu),
        #     ccf_cfDNA = ifelse(ccf_cfDNA >= 1, 1, ccf_cfDNA)
        # ) %>%
        filter(immunogenic == "yes") %>% distinct(sample, gene_name, .keep_all = TRUE),
    aes(x = ccf_FrTu, y = ccf_cfDNA, color = dna_presence, shape = sample)
) +
    geom_beeswarm(size = 3) +
    scale_color_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
    geom_text_repel(aes(label = gene_name),
        color = "grey30", size = 3.3,
        min.segment.length = unit(0, "lines"), segment.color = "grey",
        #   nudge_x = -15000,
        #   hjust = 0,
        #   direction = "y",
        max.overlaps = Inf,
        box.padding = 0.5,
        fontface = "italic", force = 100, max.time = 50
    ) +
    theme_classic() +
    labs(color = "Variant presence", shape = "Patient", x = "Mutation copy number (Tumor Biopsy)", y = "Mutation copy number (cfDNA)")

p
# ggplotly(p) %>% layout(width = 1000, height = 600)

ggsave("Immunogenic_CCF_uncapped_with_patient_name.pdf", width = 7.6, height = 6)
```

#### Filtered parameters

##### Called
- TVAF >= 3%
- Variant reads >= 4 reads
- Coverage = or > 9 in both tumor and normal samples (all 3)
- called by 2 or more callers both SNV and INDELS
- All mutations that normal VAF = 0
- If normal has VAF, VAF > 5X in tumor compared to normal

##### Present
- TVAF >= 1%

<br>
<br>

```{r}
# TO DO: change formula distinct... to include only immunogenic mutations (need to be done in CCF_calculation.Rmd files)
filter_summary_immunogenic <- ccf_table %>%
    filter(immunogenic == "yes") %>%
    distinct(sample, gene_name, .keep_all = TRUE) %>%
    mutate(
        "FrTu tvaf/nvaf" = round(vaf_FrTu / vaf_Normal, 2),
        "cfDNA tvaf/nvaf" = round(vaf_cfDNA / vaf_Normal, 2),
        vaf_FrTu = round(vaf_FrTu, 3),
        vaf_cfDNA = round(vaf_cfDNA, 3),
        vaf_Normal = round(vaf_Normal, 3),
        called_in = found_in,
        present_in = dna_presence
    ) %>%
    dplyr::select(
        sample,
        gene_name,
        `AA change`,
        called_in,
        present_in,
        ccf_FrTu,
        ccf_cfDNA,
        vaf_FrTu,
        vaf_cfDNA,
        vaf_Normal,
        `FrTu tvaf/nvaf`,
        `cfDNA tvaf/nvaf`,
        ALT_counts_FrTu,
        ALT_counts_cfDNA,
        cov_FrTu,
        cov_cfDNA,
        cov_Normal
    )

datatable(filter_summary_immunogenic)

write.table(filter_summary_immunogenic, file = "summary_neoantigens.tsv", quote = F, row.names = F, sep = "\t")
```

```{r}
#Plots summary immunogenic
# Venn diagram
FrTu_imm <- filter_summary_immunogenic %>%
    filter(present_in %in% c("FrTu_only", "Shared")) %>%
    pull(gene_name)
cfDNA_imm <- filter_summary_immunogenic %>%
    filter(present_in %in% c("cfDNA_only", "Shared")) %>%
    pull(gene_name)

x <- list(FrTu = FrTu_imm, cfDNA = cfDNA_imm)

venn <- Venn(x)
d <- process_data(venn, shape_id = "201")

ggplot() +
    geom_polygon(aes(X, Y, fill = name, group = id),
        color = "grey50",
        data = venn_regionedge(d)
    ) +
    scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
    geom_text(aes(X, Y, label = name),
        data = venn_setlabel(d)
    ) +
    # 4. region label layer
    geom_label(aes(X, Y, label = count),
        alpha = 0,
        data = venn_regionlabel(d)
    ) +
    coord_equal() +
    theme_void() +
    labs(fill = "Variant presence")
```

```{r}
ggsave("Venn_variant_presence.pdf")
```

```{r, fig.width=6.5, fig.height=3}
# Barplots mutation sources
plot_table <- filter_summary_immunogenic %>% group_by(sample)

plot_table <- filter_summary_immunogenic %>%
    group_by(sample, present_in) %>%
    summarise(n_mutations = n())

plot_table <- rbind(plot_table, data.frame(
    sample = c("GOI-02", "GOI-07"),
    present_in = c("Shared", "Shared"),
    n_mutations = c(0, 0)
))

# datatable(plot_table)

p1 <- ggplot(plot_table, aes(x = sample, y = n_mutations, fill = present_in)) +
    geom_bar(stat = "identity") +
    theme_classic() +
    scale_fill_manual(values = c("#FCC557", "#D9D8D4", "#7CC6BF")) +
    coord_flip() +
    theme(
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank()
    ) +
    labs(fill = "Variant presence", y = "# Neoantigens / patient")

# Plot % shared
ccf_summary_percent <- ccf_table %>%
    group_by(sample, dna_presence, coding_consequence) %>%
    summarise(n_mutations = n(), .groups = "drop") %>%
    group_by(sample, coding_consequence) %>%
    mutate(percent = round(n_mutations / sum(n_mutations) * 100, 2)) %>%
    ungroup()

plot_table_percent <- ccf_summary_percent %>%
    filter(coding_consequence == "non_synonymous" & dna_presence == "Shared") %>%
    mutate(square = 1)

p2 <- ggplot(plot_table_percent, aes(x = sample, y = square, color = percent)) +
    geom_point(size = 12) +
    scale_color_paletteer_c("ggthemes::Classic Gray") +
    coord_flip() +
    theme_classic() +
    theme(
        legend.position = "left",
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank()
    ) +
    labs(color = "% Shared NSMs")


# Patchwork
p2 + p1 +
    plot_layout(widths = c(1, 5))
```

```{r}
ggsave("Summary_variant_presence_barplot.pdf", height = 4, width = 7.5)
```